---
title: "Julia - US Car Accidents Analysis"
format: revealjs
editor: visual
---

# Import Data and Clean

```{r}
#install.packages("readr")
library(readr)
Data_Set_4 <- read_csv("Data_Set_4.csv")

# View column names
names(Data_Set_4)

# View unique weather conditions
Data_Set_4 |> 
  dplyr::pull(Weather_Condition) |> 
  unique()
```

# Re-Categorize Weather Conditions

There are over 100+ different weather conditions. We condensed them into 10 broader groups.

```{r}
#install.packages("dplyr")
library(dplyr)

Data_Set_4 <- Data_Set_4 |> 
  mutate(Weather_Category = case_when(
    Weather_Condition %in% c("Clear", "Fair", "Fair / Windy", "Partly Cloudy", "Mostly Cloudy", "Partly Cloudy / Windy", "Mostly Cloudy / Windy") ~ "Clear / Fair",
    Weather_Condition %in% c("Cloudy", "Overcast", "Cloudy / Windy", "Scattered Clouds") ~ "Cloudy / Overcast",
    Weather_Condition %in% c("Fog", "Mist", "Haze", "Patches of Fog", "Shallow Fog", "Light Fog", "Light Haze", "Fog / Windy", "Drizzle and Fog") ~ "Fog / Haze / Mist",
    Weather_Condition %in% c("Light Rain", "Rain", "Heavy Rain", "Drizzle", "Heavy Drizzle", "Light Rain Showers", "Rain Showers", "Rain Shower", "Showers in the Vicinity") ~ "Rain / Drizzle",
    Weather_Condition %in% c("Light Snow", "Snow", "Heavy Snow", "Snow Showers", "Blowing Snow", "Sleet", "Snow and Sleet", "Snow Grains") ~ "Snow / Sleet / Ice",
    Weather_Condition %in% c("Thunderstorm", "T-Storm", "Thunder in the Vicinity", "Thunderstorms and Rain", "Thunderstorms and Snow") ~ "Thunderstorms / Lightning",
    Weather_Condition %in% c("Blowing Dust", "Widespread Dust", "Sand", "Dust Whirls", "Squalls") ~ "Wind / Dust / Sand",
    Weather_Condition %in% c("Freezing Rain", "Freezing Drizzle", "Ice Pellets") ~ "Freezing Rain / Ice Pellets",
    Weather_Condition %in% c("Funnel Cloud", "Hail", "Small Hail", "Volcanic Ash") ~ "Other / Rare",
    TRUE ~ "Unknown"
  ))

# View updated Weather Categories
unique(Data_Set_4$Weather_Category)
```

# Car Accident Frequency by Weather Conditions

```{r}
#install.packages("ggplot2")
#install.packages("scales")
library(ggplot2)
library(scales)

Data_Set_4 |> 
  count(Weather_Category) |> 
  ggplot(aes(x = reorder(Weather_Category, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Car Accidents by Weather Category",
    x = "Weather Category",
    y = "Number of Accidents"
  ) +
  theme_minimal()
```

# Conditions and Severity

```{r}
Data_Set_4 |> 
  count(Weather_Category, Severity) |> 
  ggplot(aes(x = reorder(Weather_Category, n), y = n, fill = factor(Severity))) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Accidents by Weather and Severity",
    x = "Weather Category",
    y = "Number of Accidents",
    fill = "Severity Level"
  ) +
  theme_minimal()
```

# Re-Group Weather Categories

```{r}
Data_Set_4_grouped <- Data_Set_4 |> 
  mutate(Weather_Group = case_when(
    grepl("Rain|Drizzle", Weather_Condition, ignore.case = TRUE) ~ "Rain",
    grepl("Snow|Sleet|Ice", Weather_Condition, ignore.case = TRUE) ~ "Snow/Ice",
    grepl("Fog|Mist|Haze", Weather_Condition, ignore.case = TRUE) ~ "Fog",
    grepl("Clear|Fair", Weather_Condition, ignore.case = TRUE) ~ "Clear",
    grepl("Cloud|Overcast", Weather_Condition, ignore.case = TRUE) ~ "Cloudy",
    TRUE ~ "Other"
  ))
```

# Accident Counts by Weather Group

```{r}
Data_Set_4_grouped |> 
  count(Weather_Group) |> 
  ggplot(aes(x = reorder(Weather_Group, n), y = n)) +
  geom_col(fill = "lightgreen") +
  coord_flip() +
  labs(
    title = "Accident Counts by Weather Group",
    x = "Weather Group",
    y = "Number of Accidents"
  ) +
  theme_minimal()
```

# Chi-Square Test

```{r}
# Chi-Square Test
weather_group_counts <- Data_Set_4_grouped |> 
  count(Weather_Group)

chisq_test_result <- chisq.test(weather_group_counts$n)
chisq_test_result
```

**Interpretation:** - X² = 1,224,743 with 5 degrees of freedom. - p-value \< 2.2e-16. - We reject the null hypothesis. - Accident frequency significantly depends on weather group.

# Standardized Residuals

```{r}
chisq_test_result$stdres
```

**Interpretation:** - Cloudy and Clear: way more accidents than expected. - Rain, Other, Snow/Ice, Fog: fewer accidents than expected. - Residuals \> 2 in magnitude → highly statistically significant differences.

# Logistic Regression: Probability of Severe Accident

```{r}
# Create Binary Outcome
Data_Set_4_grouped <- Data_Set_4_grouped |> 
  mutate(Severe_Accident = ifelse(Severity >= 3, 1, 0))

# Logistic Regression Model
model <- glm(Severe_Accident ~ Weather_Group, data = Data_Set_4_grouped, family = "binomial")

# Model Summary
summary(model)
```

**Interpretation:** - All weather groups show higher odds of severe accidents compared to Clear. - Cloudy, Rain, Snow/Ice have stronger effects. - All coefficients are statistically significant (p \< 0.001).

# Key Takeaways

-   Weather conditions **significantly impact accident frequency**.
-   Clear and Cloudy days have **more accidents** mainly due to exposure.
-   Rain, Snow/Ice, and Fog have **higher per-event risk**.
-   Logistic regression shows **Cloudy, Rain, and Snow/Ice** increase odds of severe accidents.
-   Future work: Normalize by days of exposure to get true accident rates per weather type.
