---
title: "US Car Accidents"
subtitle: "Revealjs Presentation"
author: "Group 3: Anthony Fitch, Adib Mazloom, Annabelle Petherbridge, Julia Bautista, & William Jackson"
institute: "GBA 5910, Cal Poly Pomona"
date: "May 1, 2025"
format: 
  revealjs:
    theme: solarized
    width: 1600
    height: 900
    transition: slide
    transition-speed: default
    incremental: false
    toc: true
    toc-depth: 1
    slide-number: true
    scrollable: true
    smaller: true
    code-fold: false
    code-overflow: wrap
    number-sections: false
    number-depth: 5
    embed-resources: true
    css: styles.css
editor: visual
execute: 
  echo: true
  freeze: auto
---

# **RO2: HOW DOES WEATHER CONDITIONS INFLUENCE** **ACCIDENT FREQUENCY AND SEVERITY?**

```{r, include=FALSE}
#install.packages("readr")
library(readr)
Data_Set_4 <- read_csv("Data_Set_4_sample.csv")



# View unique weather conditions
Data_Set_4 |> 
  dplyr::pull(Weather_Condition) |> 
  unique()
```

## Re-Categorize Weather Conditions

There are over 100+ different weather conditions. We condensed them into 10 groups.

```{r, echo=FALSE}
#install.packages("dplyr")
library(dplyr)

Data_Set_4 <- Data_Set_4 |> 
  mutate(Weather_Category = case_when(
    Weather_Condition %in% c("Clear", "Fair", "Fair / Windy", "Partly Cloudy", "Mostly Cloudy", "Partly Cloudy / Windy", "Mostly Cloudy / Windy") ~ "Clear / Fair",
    Weather_Condition %in% c("Cloudy", "Overcast", "Cloudy / Windy", "Scattered Clouds") ~ "Cloudy / Overcast",
    Weather_Condition %in% c("Fog", "Mist", "Haze", "Patches of Fog", "Shallow Fog", "Light Fog", "Light Haze", "Fog / Windy", "Drizzle and Fog") ~ "Fog / Haze / Mist",
    Weather_Condition %in% c("Light Rain", "Rain", "Heavy Rain", "Drizzle", "Heavy Drizzle", "Light Rain Showers", "Rain Showers", "Rain Shower", "Showers in the Vicinity") ~ "Rain / Drizzle",
    Weather_Condition %in% c("Light Snow", "Snow", "Heavy Snow", "Snow Showers", "Blowing Snow", "Sleet", "Snow and Sleet", "Snow Grains") ~ "Snow / Sleet / Ice",
    Weather_Condition %in% c("Thunderstorm", "T-Storm", "Thunder in the Vicinity", "Thunderstorms and Rain", "Thunderstorms and Snow") ~ "Thunderstorms / Lightning",
    Weather_Condition %in% c("Blowing Dust", "Widespread Dust", "Sand", "Dust Whirls", "Squalls") ~ "Wind / Dust / Sand",
    Weather_Condition %in% c("Freezing Rain", "Freezing Drizzle", "Ice Pellets") ~ "Freezing Rain / Ice Pellets",
    Weather_Condition %in% c("Funnel Cloud", "Hail", "Small Hail", "Volcanic Ash") ~ "Other / Rare",
    TRUE ~ "Unknown"
  ))

# View updated Weather Categories
unique(Data_Set_4$Weather_Category)
```

```{r car-accidents-chart, include=FALSE}
#NOT INCLUDED, As 10 groups were too broad
#install.packages("ggplot2")
#install.packages("scales")
#install.packages("dplyr")
library(ggplot2)
library(scales)
library(dplyr)

Data_Set_4 |>
  count(Weather_Category) |>
  ggplot(aes(x = reorder(Weather_Category, n), y = n, fill = Weather_Category)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(
    values = c(
      "Clear / Fair" = "#FFD700",   # yellow
      "Cloudy" = "#FFA500",         # orange
      "Rain" = "#FF0000",           # red
      "Snow" = "#0000FF",           # blue
      "Fog / Haze / Other" = "#008000" # green
    )
  ) +
  labs(
    title = "Car Accidents by Weather Category",
    x = "Weather Category",
    y = "Number of Accidents",
    fill = "Weather"
  ) +
  theme_minimal()
```

## Conditions and Severity

```{r car-accident-severity-chart, echo=FALSE}
library(ggplot2)
library(scales)
library(dplyr)

# Calculate percentages and prepare labels
plot_data <- Data_Set_4 |>
  count(Weather_Category, Severity) |>
  group_by(Weather_Category) |>
  mutate(
    pct = n / sum(n),
    label = ifelse(pct >= 0.10, percent(pct, accuracy = 1), "")  # Only show if ≥5%
  ) |>
  ungroup()

# Plot with improved label styling
ggplot(plot_data, aes(x = reorder(Weather_Category, n), y = n, fill = factor(Severity))) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 4,                # Slightly bigger text
    fontface = "bold",       # Make it bold
    color = "white"          # White text for contrast
  ) +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(
    values = c(
      "1" = "#FFD700",   # yellow
      "2" = "#FFA500",   # orange
      "3" = "#FF0000",   # red
      "4" = "#0000FF",   # blue
      "5" = "#008000"    # green (if applicable)
    )
  ) +
  labs(
    title = "Accidents by Weather and Severity",
    x = "Weather Category",
    y = "Number of Accidents",
    fill = "Severity Level"
  ) +
  theme_minimal(base_size = 14)

```

-   **Most accidents under all weather conditions are classified as Severity Level 2 (orange)**, indicating moderate severity is the most common outcome regardless of weather type.

-   **Severe accidents (Levels 3 and 4) occur more frequently during Clear / Fair conditions** than in adverse weather, likely due to higher traffic volumes or driver overconfidence.

## Re-Group Weather Categories

```{r}
Data_Set_4_grouped <- Data_Set_4 |> 
  mutate(Weather_Group = case_when(
    grepl("Rain|Drizzle", Weather_Condition, ignore.case = TRUE) ~ "Rain",
    grepl("Snow|Sleet|Ice", Weather_Condition, ignore.case = TRUE) ~ "Snow/Ice",
    grepl("Fog|Mist|Haze", Weather_Condition, ignore.case = TRUE) ~ "Fog",
    grepl("Clear|Fair", Weather_Condition, ignore.case = TRUE) ~ "Clear",
    grepl("Cloud|Overcast", Weather_Condition, ignore.case = TRUE) ~ "Cloudy",
    TRUE ~ "Other"
  ))
```

## Accident Counts by Weather Group

```{r,echo=FALSE}
plot_data <- Data_Set_4_grouped |>
  count(Weather_Group, Severity) |>
  group_by(Weather_Group) |>
  mutate(
    Total = sum(n),
    Percent = n / Total * 100
  ) |>
  ungroup()

# Plot
ggplot(plot_data, aes(x = reorder(Weather_Group, Total), y = n, fill = factor(Severity))) +
  geom_col(position = "stack") +
geom_text(
  aes(label = ifelse(Percent >= 20, paste0(round(Percent, 0), "%"), "")),
  position = position_stack(vjust = 0.5),
  size = 3,
  color = "black",
  family = "sans"
) +
  coord_flip() +
  scale_y_continuous(labels = comma) +  # comma-formatted y-axis
  labs(
    title = "Accident Counts by Weather Group with Severity % Labels",
    x = "Weather Condition",
    y = "Number of Accidents",
    fill = "Severity Level"
  ) +
  theme_minimal(base_size = 14) +  # larger base font
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11)
  )
```

-   **Clear and Cloudy conditions** have the highest overall number of accidents by far compared to other weather groups.

-   Across all weather types, **Severity Level 2 (likely moderate accidents)** makes up the largest portion—typically around **75–80%**.

-   In **Cloudy and Clear weather**, about **20–21%** of accidents are more serious (Severity Level 3), which is higher than in other conditions.

-   **Rain, Fog, Snow/Ice, and Other** weather groups show relatively **few severe accidents** above 25%, so only the most frequent severity is labeled.

## Chi-Square Test

```{r,echo=FALSE}
# Chi-Square Test
weather_group_counts <- Data_Set_4_grouped |> 
  count(Weather_Group)

chisq_test_result <- chisq.test(weather_group_counts$n)
chisq_test_result
```

-   This Chi-Square test was conducted on a **random sample of 10,000 rows** from the full dataset.

-   The **p-value is extremely small** (p \< 2.2e-16), suggesting that the result is **statistically significant**.

-   We **reject the null hypothesis**, which assumes accident counts are evenly distributed across weather groups.

-   This implies that **accident frequency significantly varies by weather group**—some weather conditions are associated with more accidents.

-   Since the test was based on a **sample**, results may vary slightly across samples, but the **strong significance suggests a real relationship** if the sample is representative of the population.

## Standardized Residuals

```{r}
chisq_test_result$stdres
```

-   Standardized residuals were calculated from a **random 10,000-row sample** of the full dataset.

-   **Cloudy and Clear** weather categories had **much more accidents than expected**.

-   **Rain, Fog, Snow/Ice, and Other** had **fewer accidents than expected** based on the sample.

-   All residuals are **well beyond ±2**, which suggests the differences are **highly statistically significant**.

-   These insights are valid **for the sample**, and likely generalize if the sample is representative of the full dataset.

## Logistic Regression: Probability of Severe Accident

```{r, echo=FALSE}
# Create Binary Outcome
Data_Set_4_grouped <- Data_Set_4_grouped |> 
  mutate(Severe_Accident = ifelse(Severity >= 3, 1, 0))

# Logistic Regression Model
model <- glm(Severe_Accident ~ Weather_Group, data = Data_Set_4_grouped, family = "binomial")

# Model Summary
summary(model)
```

#### Why Use Logistic Regression?

Logistic regression is used to model **binary outcomes**—in this case, whether an accident is **severe (1)** or **not severe (0)**.

-   It estimates the **probability** of a severe accident under different **weather conditions**.

-   The model compares each weather type to a **baseline (Clear)** to determine which ones **significantly affect risk**.

#### Key Takeaways (Based on a Sample of 10,000 Rows):

-   **Clear weather is the baseline**: All other weather types are compared to it.

-   **Bad weather = higher risk**: Cloudy, foggy, and rainy conditions increase the probability of a severe accident compared to clear weather.

-   **Cloudy and rainy weather are especially risky**, with the largest and most statistically significant coefficients.

-   **Most coefficients are statistically significant (p \< 0.001)**, meaning the results are unlikely due to chance.

# Key Takeaways

**Weather significantly influences accidents** — frequency and severity vary across conditions (*p* \< 0.001) in the sample analysis.

**Clear and Cloudy weather account for the most accidents**, likely due to higher traffic volume and exposure.

**Severity Level 2 (moderate)** is the most common outcome across all weather types.

**Cloudy and Rainy conditions increase the odds of severe accidents**, as shown in the logistic regression model.
