---
title: "Anthony Fitch"
subtitle: "Revealjs Presentation"
author: "Anthony Fitch"
institute: "Your Course Number, Cal Poly Pomona"
date: "`r Sys.Date()`"
format: 
  revealjs:
    theme: serif
    width: 1600
    height: 900
    footer: Title of Your Project
    transition: slide
    transition-speed: default
    incremental: false
    toc: true
    toc-depth: 1
    slide-number: true
    scrollable: true
    smaller: true
    code-fold: false
    code-overflow: wrap
    number-sections: false
    number-depth: 5
    embed-resources: true
    css: styles.css
editor: visual
execute: 
  echo: true
  freeze: auto
---

## Data Wrangling 2017 Census Data

```{r}

library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)

Census_Data_1 <- read.csv("2017 Census Data.csv")

dim(Census_Data_1)

#names(Census_Data_1)

```

## Column Names

\[1\] "TractId" "State" "County" "TotalPop" \[5\] "Men" "Women" "Hispanic" "White" \[9\] "Black" "Native" "Asian" "Pacific" \[13\] "VotingAgeCitizen" "Income" "IncomeErr" "IncomePerCap" \[17\] "IncomePerCapErr" "Poverty" "ChildPoverty" "Professional" \[21\] "Service" "Office" "Construction" "Production" \[25\] "Drive" "Carpool" "Transit" "Walk" \[29\] "OtherTransp" "WorkAtHome" "MeanCommute" "Employed" \[33\] "PrivateWork" "PublicWork" "SelfEmployed" "FamilyWork" \[37\] "Unemployment"

## Top 20 States: Men Vs. Women Population

```{r}
#unique(Census_Data_1$State)

# 1. Prepare the data
top_states_long <- Census_Data_1 %>%
  group_by(State) %>%
  summarise(
    Total_Men = sum(Men, na.rm = TRUE),
    Total_Women = sum(Women, na.rm = TRUE)
  ) %>%
  mutate(TotalPop = Total_Men + Total_Women) %>%
  arrange(desc(TotalPop)) %>%
  slice_head(n = 20) %>%
  pivot_longer(cols = c(Total_Men, Total_Women),
               names_to = "Gender",
               values_to = "Population") %>%
  group_by(State) %>%
  mutate(Percent = Population / sum(Population) * 100)

# 2. Plot
ggplot(top_states_long, aes(x = reorder(State, TotalPop), y = Population, fill = Gender)) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = paste0(round(Percent, 1), "%")),
    position = position_stack(vjust = 0.5),
    color = "white",
    size = 3
  ) +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = c("Total_Men" = "red", "Total_Women" = "purple")) +
  labs(
    title = "Top 20 States by Population (Men vs Women Stacked with %)",
    x = "State",
    y = "Population",
    fill = "Gender"
  ) +
  theme_minimal()


```

## Wrangling US Education Data

```{r}
US_Education <- read.csv("US_Educational_Data.csv")

dim(US_Education)

names(US_Education)

library(dplyr)
library(tidyr)

# Step 1: Create a new column called County
US_Education <- US_Education %>%
  mutate(County = ifelse(grepl("County", Label..Grouping.), Label..Grouping., NA)) %>%
  fill(County)

# Step 2 (optional): If you want, you can also clean the County name (remove extra stuff if needed)
# For example, remove extra text like "County:" if it's messy
# US_Education <- US_Education %>%
#   mutate(County = str_remove(County, "County: "))  # optional cleaning

US_Education <- US_Education %>%
  relocate(County, .before = 1)

print(US_Education)

library(googledrive)

#url <- "https://drive.google.com/file/d/1tukQ6qwB9v4BbKg8oY_ECVBXilDcAnWP/view?usp=share_link"
#data <- read.csv(url)


```

## Data Wrangling US Car Accident Data
