---
title: "US Car Accidents"
subtitle: "Revealjs Presentation"
author: "Group 3: Anthony Fitch, Adib Mazloom, Annabelle Petherbridge, Julia Bautista, & William Jackson"
institute: "GBA 5910, Cal Poly Pomona"
date: "May 1, 2025"
format: 
  revealjs:
    theme: solarized
    width: 1600
    height: 900
    transition: slide
    transition-speed: default
    incremental: false
    toc: true
    toc-depth: 1
    slide-number: true
    scrollable: true
    smaller: true
    code-fold: true
    code-overflow: wrap
    number-sections: false
    number-depth: 5
    embed-resources: true
    css: styles.css
editor: visual
execute: 
  echo: true
  freeze: auto
---

# Research Objective #1

What counties have the most accidents per-capita and overall. Compare these counties across multiple characteristics, age, income, etc. using US census Data.

## Dependent Variable

-   Accidents Per Capita

-   I am analyzing this against various Independent Variables \## Dependent Variables

-   Commuting Behavior

    -   Walk
    -   Drive
    -   Carpool
    -   Transit
    -   Work at Home
    -   Mean Commute

-   Employment Type

    -   Professional
    -   Service
    -   Office

## Operationalization of the Scale

-   **Dependent Variables**
-   Accidents Per Capita
    -   Operationalization
        -   Total Accidents per county/ Total Population of County
    -   Scale Type
        -   Ratio
-   **Independent Variables**

| Variable     | Operationalization                | Scale Type |
|--------------|-----------------------------------|------------|
| Walk         | \% of workers who walk to work    | Ratio      |
| Drive        | \% who drive alone                | Ratio      |
| Carpool      | \% who carpool                    | Ratio      |
| Transit      | \% who use public transit         | Ratio      |
| WorkAtHome   | \% who work remotely              | Ratio      |
| MeanCommute  | Average commute time in minutes   | Ratio      |
| Professional | %employed in professional roles   | Ratio      |
| Service      | \% employed in service jobs       | Ratio      |
| Office       | \% employed in office/admin roles | Ratio      |
|              |                                   |            |

## Data Wrangling

-   Refer to: Refernce data, of bigger data sets
    -   US Traffic Accident
        -   Data had 7 million rows so I wrangled the data and only kept sums and tables I needed
    -   US Census Data
        -   Originally had 100,000 rows and wrangled it down to 14,000 rows which is all i needed based on the counties i found in the US traffic Accident data

```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(scales)
library(patchwork)
library(tidytext)

#Read the big dataset with 7 million rows
#US_Accident_Data <- read_csv("US_Accidents.csv")

#Read Census data
#US_Census_Data <- read_csv("2017 Census Data.csv")



# Summarize and arrange
#top_counties <- US_Accident_Data %>%
 # count(County, sort = TRUE) %>%
  #top_n(20, n)

# Plot
#ggplot(top_counties, aes(x = reorder(County, n), y = n)) +
  #geom_bar(stat = "identity") +
  #coord_flip() +
  #labs(title = "Top 20 Counties by Accident Count", x = "County", y = "Count")

#top_counties_df <- top_counties %>%
  #rename(Accident_Count = n)

#print(top_counties_df)

#write.csv(top_counties_df, "top_20_counties.csv", row.names = FALSE)

Top_20_Counties <- read_csv("top_20_counties.csv")

print(Top_20_Counties)


```

## Data Wrangling: Plot of top 20 counties with the most accidents

```{r}
ggplot(Top_20_Counties, aes(x = reorder(County, Accident_Count), y = Accident_Count)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = Accident_Count), hjust = -0.1, size = 3) +
  coord_flip() +
  labs(
    title = "Top 20 Counties by Accident Count",
    x = "County",
    y = "Accident Count"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  # Center title
```

## Data Wrangling: US Census

-   Check for Missing or Incosistent data
    -   None found
-   Standardize columns with ratios as percentages
-   Data was already Tidy
-   Data was already long

```{r}
US_Census_Counties <- read_csv("US_Census_Counties.csv")

#Check for missing or inconsistent data
summary(US_Census_Counties)
colSums(is.na(US_Census_Counties))

#US_Census_Counties <- US_Census_Counties %>%
  #mutate(across(c(Walk, Drive, Carpool, Transit, WorkAtHome), ~ ./100))

County_Pop_Summary <- US_Census_Counties %>%
  group_by(County) %>%
  summarise(Total_Population = sum(TotalPop, na.rm = TRUE)) %>%
  arrange(desc(Total_Population))


```

## Data Wrangling: Finding top County Populations

```{r}
ggplot(County_Pop_Summary, aes(x = reorder(County, Total_Population), y = Total_Population)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  scale_y_continuous(labels = comma) +  # Use commas, not scientific notation
  labs(
    title = "Total Population by County",
    x = "County",
    y = "Total Population"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

## Data Wrangling: Merging Datasets Total Accidents and Total Population

```{r}
Top_20_Counties_with_Pop <- Top_20_Counties %>%
  left_join(County_Pop_Summary %>% select(County, Total_Population), by = "County")

Top_20_Counties_with_Pop <- Top_20_Counties_with_Pop %>%
  mutate(Accidents_Per_Capita = Accident_Count / Total_Population)

print(Top_20_Counties_with_Pop)



```

## Variable: Graph Showing Total Accident Counts per County

```{r}

top_15_count <- Top_20_Counties_with_Pop %>%
  arrange(desc(Accident_Count)) %>%
  slice_head(n = 15) %>%
  mutate(County = factor(County, levels = rev(County)))  # fix order

plot1 <- ggplot(top_15_count, aes(x = County, y = Accident_Count, fill = County)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 15 Counties by Total Accident Count",
    x = "County",
    y = "Accident Count"
  ) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
print(plot1)
```

## Variable: Showing Per Capita which of the Top Counties are the worst

```{r}
top_15_per_capita <- Top_20_Counties_with_Pop %>%
  arrange(desc(Accidents_Per_Capita)) %>%
  slice_head(n = 15) %>%
  mutate(County = factor(County, levels = rev(County)))  # fix order

plot2 <- ggplot(top_15_per_capita, aes(x = County, y = Accidents_Per_Capita, fill = County)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 15 Counties by Accidents Per Capita",
    x = "County",
    y = "Accidents per Capita"
  ) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

print(plot2)
```

## Data Visualization: Graph of Per Capita and % of Walkers

```{r}
# Step 1: Define counties of interest
selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Step 2: Walk data
walk_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(Walk, na.rm = TRUE)) %>%
  mutate(Metric = "Average Walk (%)")

# Step 3: Accident data (this determines the order)
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

# Step 4: Set county order based on descending Accidents per Capita
county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

# Step 5: Combine both and apply consistent order
combined_df <- bind_rows(walk_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))  # reverse for coord_flip()

# Step 6: Plot
ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Walk Percentage",
    x = "County",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold", size = 12)
  )
```

## Data Visuzliation: Graph of Per Capita and % Work From Home

```{r}
# Step 1: Define selected counties
selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Step 2: Work-from-home data
wfh_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(WorkAtHome, na.rm = TRUE)) %>%
  mutate(Metric = "Average Work-from-Home (%)")

# Step 3: Accident per capita data (for ordering)
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

# Step 4: Define County order based on descending Accidents per Capita
county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

# Step 5: Combine datasets and apply consistent factor order
combined_df <- bind_rows(wfh_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))  # reverse for coord_flip

# Step 6: Create side-by-side plot
ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Work-from-Home Percentage",
    x = "County",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold", size = 12)
  )
```

## Data Visualization: Graph with Per capita and Carpool

```{r}
# Step 1: Define selected counties
selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Step 2: Carpool data
carpool_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(Carpool, na.rm = TRUE)) %>%
  mutate(Metric = "Average Carpool (%)")

# Step 3: Accident per capita data (used for ordering)
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

# Step 4: Define County order by descending Accidents per Capita
county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

# Step 5: Combine both datasets and enforce consistent County order
combined_df <- bind_rows(carpool_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))  # reversed for coord_flip()

# Step 6: Create faceted plot
ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Carpool Percentage",
    x = "County",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold", size = 12)
  )
```

## Data Visuzliation: Graph with per Capita and % of Professional Jobs

```{r}
# Step 1: Define selected counties
selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Step 2: Professional employment data
professional_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(Professional, na.rm = TRUE)) %>%
  mutate(Metric = "Average Professional Employment (%)")

# Step 3: Accident per capita data (used for ordering)
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

# Step 4: Set County order based on descending Accidents per Capita
county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

# Step 5: Combine both datasets with shared County factor levels
combined_df <- bind_rows(professional_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))  # reversed for coord_flip()

# Step 6: Create the side-by-side faceted plot
ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Professional Employment",
    x = "County",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold", size = 12)
  )
```

## Data Visualization: Graph of Per Capita and Service Jobs

```{r}
# Step 1: Define selected counties
selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Step 2: Service employment data
service_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(Service, na.rm = TRUE)) %>%
  mutate(Metric = "Average Service Employment (%)")

# Step 3: Accidents per capita data (used for ordering)
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

# Step 4: Set County order based on descending Accidents per Capita
county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

# Step 5: Combine both datasets and apply the same county order
combined_df <- bind_rows(service_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))  # reversed for coord_flip()

# Step 6: Create the side-by-side faceted plot
ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Service Employment",
    x = "County",
    y = "Value"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold", size = 12)
  )
```

## Data Visualization: Graph of Per Capita and % Office Employment

```{r}
#Office

selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

office_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(Office, na.rm = TRUE)) %>%
  mutate(Metric = "Average Office Employment (%)")

accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

combined_df <- bind_rows(office_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))

ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Office Employment",
    x = "County",
    y = "Value"
  ) +
  theme_minimal()
```

## Data Visualization: Graph of Per Capita and % Who Drive alone

```{r}
#Drive

drive_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(Drive, na.rm = TRUE)) %>%
  mutate(Metric = "Average Drive Alone (%)")

accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

combined_df <- bind_rows(drive_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))

ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Drive Alone",
    x = "County",
    y = "Value"
  ) +
  theme_minimal()
```

## Data Visualization: Graph of Per Capita and % Commuting with Public Transit

```{r}
#Transit

transit_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(Transit, na.rm = TRUE)) %>%
  mutate(Metric = "Average Public Transit (%)")

accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

combined_df <- bind_rows(transit_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))

ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Public Transit Usage",
    x = "County",
    y = "Value"
  ) +
  theme_minimal()
```

## Data Visualization: Graph of Per Capita and Mean Commute in Minutes

```{r}
#MeanCommute

commute_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(Value = mean(MeanCommute, na.rm = TRUE)) %>%
  mutate(Metric = "Average Commute Time (min)")

accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  transmute(County, Value = Accidents_Per_Capita, Metric = "Accidents per Capita")

county_order <- accident_df %>%
  arrange(desc(Value)) %>%
  pull(County)

combined_df <- bind_rows(commute_df, accident_df) %>%
  mutate(County = factor(County, levels = rev(county_order)))

ggplot(combined_df, aes(x = County, y = Value, fill = County)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ Metric, scales = "free_x", nrow = 1) +
  labs(
    title = "County Comparison: Accidents per Capita vs. Commute Time",
    x = "County",
    y = "Value"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and Mean Commute

```{r}
# Step 1: Join commute and accident data directly for scatter plot
selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Commute Time
commute_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MeanCommute = mean(MeanCommute, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(commute_df, accident_df, by = "County")

# Step 2: Scatter plot with regression line
ggplot(joined_df, aes(x = MeanCommute, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Mean Commute Time vs. Accidents per Capita",
    x = "Average Commute Time (min)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and Transit

```{r}
# Scatter Plot: Public Transit (%) vs. Accidents per Capita


selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Public Transit (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(Transit, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Public Transit (%) vs. Accidents per Capita",
    x = "Public Transit (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and Drive

```{r}
# Scatter Plot: Drive Alone (%) vs. Accidents per Capita


selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Drive Alone (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(Drive, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Drive Alone (%) vs. Accidents per Capita",
    x = "Drive Alone (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and Office

```{r}
# Scatter Plot: Office Employment (%) vs. Accidents per Capita


selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Office Employment (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(Office, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Office Employment (%) vs. Accidents per Capita",
    x = "Office Employment (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and Service

```{r}
# Scatter Plot: Service Employment (%) vs. Accidents per Capita


selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Service Employment (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(Service, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Service Employment (%) vs. Accidents per Capita",
    x = "Service Employment (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and Professional

```{r}
# Scatter Plot: Professional Employment (%) vs. Accidents per Capita


selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Professional Employment (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(Professional, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Professional Employment (%) vs. Accidents per Capita",
    x = "Professional Employment (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()

```

## Scatter Plot Per Capita and Carpool

```{r}
# Scatter Plot: Carpool (%) vs. Accidents per Capita


selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Carpool (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(Carpool, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Carpool (%) vs. Accidents per Capita",
    x = "Carpool (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and WorkAtHome

```{r}
# Scatter Plot: Work From Home (%) vs. Accidents per Capita

selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Work From Home (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(WorkAtHome, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Work From Home (%) vs. Accidents per Capita",
    x = "Work From Home (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Scatter Plot Per Capita and Walk

```{r}
# Scatter Plot: Walk (%) vs. Accidents per Capita


selected_counties <- c(
  "Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas", "Mecklenburg",
  "Montgomery", "Wake", "San Bernardino", "Travis", "Maricopa",
  "San Diego", "Alameda", "Davidson", "Sacramento"
)

# Metric: Walk (%)
metric_df <- US_Census_Counties %>%
  filter(County %in% selected_counties) %>%
  group_by(County) %>%
  summarise(MetricValue = mean(Walk, na.rm = TRUE))

# Accidents per Capita
accident_df <- Top_20_Counties_with_Pop %>%
  filter(County %in% selected_counties) %>%
  select(County, Accidents_Per_Capita)

# Combine
joined_df <- left_join(metric_df, accident_df, by = "County")

# Scatter Plot with Best Fit Line
ggplot(joined_df, aes(x = MetricValue, y = Accidents_Per_Capita, label = County)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred", linetype = "dashed") +
  geom_text(nudge_y = 0.001, size = 3, check_overlap = TRUE) +
  labs(
    title = "Relationship: Walk (%) vs. Accidents per Capita",
    x = "Walk (%)",
    y = "Accidents per Capita"
  ) +
  theme_minimal()
```

## Table Summary of Research Objective #1

-   In conclusion

| Variable | Summary | Accidents Per Capita | Independent Variable |
|------------------|------------------|------------------|------------------|
| Top 20 Counties by Accident Count | Los Angeles, Miami-Dade, Orange, Harris, Dallas, etc. |  |  |
| Total Population by County | Nothing of Importance |  |  |
| Top 15 Counties by Accident Count | Had missing data so reduced to top 15 |  |  |
| Top 15 Counties Accidents per Capita | This data was a better representation of how a county is doing |  |  |
| Per Capita and Mean Commute | The worst counties of accidents per capita have the lower average commute | High Rate | Lower Commute |
| Per Capita and Transit | Counties with a higher percentage of public transit have less accidents per capita | Low Rate | High Public Transit |
| Per Capita and One Driver | Accidents per Capita go up as the % of people who drive alone go up | High Rate | High amount of people driving alone |
| Per Capita and people who work in an office | No Findings | No Slope | No Slope |
| Per Capita and Service Workers | A steep decline showing a higher percentage of Service employment relates to a lower Accident rate | Low Rate | High amount of Service Workers |
| Per Capita and Professional Employment | A rise in Professional employment shows a rise in Accidents per capita | High Rate | High Professional Employment |
| Per Capita and those who Carpool | An increase in people who carpool shows a decrease in Accidents per Capita | Low Rate | High Carpool Rates |
| Per Capita and those who work from home | An increase in people working from home shows an increase in Accidents per Capita | High Rate | High amount of people working from home |
| Per Capita and those who walk to work | More people who walk to work shows a decrease in Accidents per Capita | Low Rate | High amount of people walking to work |

## Summary of Research Objective #1

-   Accidents per Capita go UP when:

    -   Lower Commute times (due to density)

    -   More people are driving alone

    -   More Professional Employment

    -   More People working from home

-   Accidents per Capita go DOWN when:

    -   Higher amount of Public Transit

    -   More Service Workers

    -   More people carpooling

    -   More people walking to work

## Reference Only

-   Refernce data, of bigger data sets
    -   US Traffic Accident
        -   Data had 7 million rows so I wrangled the data and only kept sums and tables I needed -US Census Data
        -   Originally had 100,000 rows and wrangled it down to 14,000 rows which is all i needed based on the counties i found in the US traffic Accident data

```{r}
#Read Census data
#US_Census_Data <- read_csv("2017 Census Data.csv")

#head(US_Census_Data)

# Remove the word "County" and trim any extra whitespace
#US_Census_Data$County <- gsub("County", "", US_Census_Data$County)
#US_Census_Data$County <- trimws(US_Census_Data$County)

# Clean list of counties (de-duplicated and trimmed)
#target_counties <- c(
  #"Los Angeles", "Miami-Dade", "Orange", "Harris", "Dallas",
  #"Mecklenburg", "Montgomery", "Wake", "San Bernardino", "Travis",
  #"Maricopa", "San Diego", "Alameda", "Davidson", "Sacramento",
 # "Riverside", "East Baton Rouge", "Santa Clara", "Cook", "Marion",
  #"Fulton", "Jefferson", "Washington", "Jackson", "Pima", "Greenville",
 # "El Paso"
#)

# Create a new dataset with only those counties
#filtered_census_data <- US_Census_Data %>%
 # filter(County %in% target_counties)

#print(filtered_census_data)

#write.csv(filtered_census_data, "US_Census_Counties.csv", row.names = FALSE)
```

```{r}
# Ensure Severity is treated as a factor
#US_Accident_Data$Severity <- as.factor(US_Accident_Data$Severity)

# Function to get top 10 counties for a given severity level
#get_top_counties_by_severity <- function(severity_level) {
 # US_Accident_Data %>%
    #filter(Severity == severity_level) %>%
    #count(County, sort = TRUE) %>%
    #top_n(10, n) %>%
   # arrange(desc(n)) %>%
   # rename(Accident_Count = n)
#}

# Create datasets
#top_counties_sev1 <- get_top_counties_by_severity("1")
#top_counties_sev2 <- get_top_counties_by_severity("2")
#top_counties_sev3 <- get_top_counties_by_severity("3")
#top_counties_sev4 <- get_top_counties_by_severity("4")

# Export to CSV files
#write.csv(top_counties_sev1, "top_counties_severity1.csv", row.names = FALSE)
#write.csv(top_counties_sev2, "top_counties_severity2.csv", row.names = FALSE)
#write.csv(top_counties_sev3, "top_counties_severity3.csv", row.names = FALSE)
#write.csv(top_counties_sev4, "top_counties_severity4.csv", row.names = FALSE)
```
