---
title: "Improving Road Safety and Traffic Management: Data Analysis and Visualization of Traffic Accidents"
subtitle: "Final Report"
author: "Anthony Fitch, Adib Mazloom, Annabelle Petherbridge, Julia Bautista, & William Jackson"
institute: "GBA 5910, Cal Poly Pomona"
date: "`r Sys.Date()`"
format:
  revealjs:
    theme: simple
    transition: fade
    toc: true
    slide-number: true
    smaller: true
execute:
  echo: true
  warning: false
  message: false
---

## Introduction

In this section of the group project, we will explore two primary research questions:

1. What is the most common **type** of traffic accident?
2. What are the most common **causes** of those accidents?

---

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)

# Load the dataset
accidents <- read_csv("cleaned.csv")

# View a sample of the data
head(accidents)
```

## Most Common Accident Types (Bar Chart)

```{r accident-type-bar, fig.width=8, fig.height=5, echo=FALSE}
accidents %>%
  count(Type_of_collision, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(Type_of_collision, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Most Common Accident Types",
       x = "Accident Type",
       y = "Count") +
  theme_minimal()
```

---

## Analysis: Bar Chart - Accident Types

- Rear-end collisions are the most common, followed by head-on and side-swipe crashes.
- These often result from inattention, tailgating, and failure to follow traffic rules.
- Driver’s ed can focus more on reaction time and following distance.
- Techniques like simulated braking and interactive drills can reinforce safer habits.
- AR scenarios can improve spatial awareness and lane control.

---

## Most Common Accident Types (Pie Chart)

```{r accident-type-pie, fig.width=6, fig.height=6, echo=FALSE}
accidents %>%
  count(Type_of_collision, sort = TRUE) %>%
  top_n(5) %>%
  ggplot(aes(x = "", y = n, fill = Type_of_collision)) +
  geom_col(width = 1) +
  coord_polar("y") +
  labs(title = "Accident Type Distribution") +
  theme_void()
```

---

## Analysis: Pie Chart - Accident Types

- A few types dominate most accidents: rear-end, head-on, and single-vehicle.
- Emphasizing these in training helps allocate learning time efficiently.
- Virtual driving sims could offer specific high-risk scenario training.
- Short-form video modules with real dashcam footage can boost attention and retention.

---

## Most Common Accident Causes (Bar Chart)

```{r accident-cause-bar, fig.width=8, fig.height=5, echo=FALSE}
accidents %>%
  count(Cause_of_accident, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(Cause_of_accident, n), y = n)) +
  geom_bar(stat = "identity", fill = "tomato") +
  coord_flip() +
  labs(title = "Top 10 Causes of Accidents",
       x = "Cause",
       y = "Count") +
  theme_minimal()
```

---

## Analysis: Bar Chart - Accident Causes

- Speeding and distracted driving are the top causes of accidents.
- These behaviors reduce response time and lead to serious injuries.
- Traditional lectures are not enough to address behavioral habits.
- Distraction simulators and gamified modules could engage students better.
- Real stories and interactive challenges create emotional impact.

---

## Most Common Accident Causes (Pie Chart)

```{r accident-cause-pie, fig.width=6, fig.height=6, echo=FALSE}
accidents %>%
  count(Cause_of_accident, sort = TRUE) %>%
  top_n(5) %>%
  ggplot(aes(x = "", y = n, fill = Cause_of_accident)) +
  geom_col(width = 1) +
  coord_polar("y") +
  labs(title = "Accident Cause Distribution") +
  theme_void()
```

---

## Analysis: Pie Chart - Accident Causes

- Speeding and distractions make up the majority of accidents.
- These are habit-based, not knowledge-based mistakes.
- Driver contracts and self-evaluations can build accountability.
- Apps with post-drive feedback can turn real habits into learning moments.
- AI tools can coach new drivers with personalized safety tips.

---

## Summary & Recommendations

- **Rear-end and head-on collisions** are the most common accident types.
- **Speeding and distracted driving** are the primary causes.
- Drivers' education should evolve to use:
  - Interactive and immersive technologies (e.g., AR/VR)
  - Scenario-based learning with decision trees
  - Behavior-modifying feedback tools

By targeting the most frequent accident types and causes, training programs can become more effective, data-driven, and life-saving.
