---
title: "US Car Accidents"
subtitle: "Revealjs Presentation"
author: "Group 3: Anthony Fitch, Adib Mazloom, Annabelle Petherbridge, Julia Bautista, & William Jackson"
institute: "GBA 5910, Cal Poly Pomona"
date: "May 1, 2025"
format: 
  revealjs:
    theme: solarized
    width: 1600
    height: 900
    transition: slide
    transition-speed: default
    incremental: false
    toc: true
    toc-depth: 1
    slide-number: true
    scrollable: true
    smaller: true
    code-fold: true
    code-overflow: wrap
    number-sections: false
    number-depth: 5
    embed-resources: true
    css: styles.css
editor: visual
execute: 
  echo: true
  freeze: auto
---

# Research Objective #4

This study aims to analyze the temporal distribution of car accidents, focusing on the frequency of incidents by day of the week and time of day.

**Dependent Variable:**

-   **Accident Frequency**: The number of accidents occurring within specific time intervals (e.g., hourly or daily).

**Independent Variables:**

-   **Time of Day**: Categorized into periods such as morning (6–9 AM), afternoon (12–3 PM), evening (4–7 PM), and night (8 PM–5 AM).

-   **Day of Week**: Days categorized from Monday to Sunday.

**Operationalization of Scale:**

| **Variable** | **Operationalization**                        | **Scale Type** |
|-----------------|---------------------------------------|-----------------|
| **Day**      | Day of the week (e.g., Monday, Tuesday, etc.) | Ordinal        |
| **Hour**     | Hour of the day (24-hour format)              | Ratio          |

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(plotly)
library(lubridate)

# Define the raw file URL from GitHub
url <- "https://raw.githubusercontent.com/BBautisJuju/US-Car-Accidents---Group-3/refs/heads/main/RTA%20Dataset.csv"

# Load the data directly into R
data <- read.csv(url)

# Create hour column
data <- data %>%
mutate(Hour = as.numeric(format(strptime(Time, "%H:%M:%S"), "%H")))

```

# Accidents by Day of Week

```{r, echo=FALSE}
#| label: accidents-by-day
#| echo: true
#| message: false
#| warning: false

accidents_by_day <- data %>%
  count(Day_of_week) %>%
  arrange(n) %>%
  mutate(Day_of_week = factor(Day_of_week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))

ggplot(accidents_by_day, aes(x = Day_of_week, y = n, fill = Day_of_week)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), vjust = -0.3) +
  labs(x = "Day of the Week", y = "Number of Accidents", title = "Accidents by Day of the Week") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Analysis:**

-   Counted the number of accidents for each day of the week

-   Visualized the distribution using a bar chart

**Findings:**

-   Accident count is highest on weekdays.

-   The greatest number of accidents occur on Fridays.

-   The fewest number of accidents occur on Saturday and Sunday.

# Accidents by Hour

```{r, echo=FALSE}
#| label: accidents-by-time
#| echo: true
#| message: false
#| warning: false

# Count accidents by hour
accidents_by_hour <- data %>%
  count(Hour) %>%
  arrange(Hour)

# Plot
ggplot(accidents_by_hour, aes(x = Hour, y = n)) +
  geom_line(group = 1, color = "darkred") +
  geom_point(color = "darkred") +
  geom_text(aes(label = n), vjust = -0.3, size = 3.5) +
  labs(x = "Hour of the Day", y = "Number of Accidents", title = "Accidents by Time of Day") +
  theme_minimal()
```

**Analysis:**

-   Counted the number of accidents for each hour of the day

-   Visualized the distribution using a line chart

**Findings:**

-   Accident frequency is highest at 8am and 5pm.

-   **Morning Rush (8:00 AM)**: This peak coincides with the morning commute, where individuals are traveling to work or school. The roads are more congested, leading to increased chances of accidents

-   **Evening Rush (5:00 PM):** The evening peak aligns with the end of the workday, leading to another surge in traffic as people head home.

# Accident Frequency: Weekday vs Weekend

```{r, echo=FALSE}
#| label: weekday-vs-weekend-overall
#| echo: true
#| message: false
#| warning: false

# Create 'Weekend' variable
data <- data %>%
  mutate(Weekend = ifelse(Day_of_week %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))

# Count accidents by 'Weekend' and 'Hour'
accidents_by_type_time <- data %>%
  count(Weekend, Hour) %>%
  arrange(Weekend, Hour)

# Calculate overall accidents by 'Hour'
accidents_overall <- data %>%
  count(Hour) %>%
  mutate(Weekend = "Overall") %>%
  arrange(Hour)

# Combine
combined_data <- bind_rows(accidents_by_type_time, accidents_overall)

# Plot
ggplot(combined_data, aes(x = Hour, y = n, color = Weekend)) +
  geom_line() +
  labs(x = "Hour of the Day", y = "Number of Accidents", title = "Accidents by Time of Day: Weekday, Weekend, and Overall") +
  scale_color_manual(values = c("Weekday" = "blue", "Weekend" = "red", "Overall" = "black")) +
  theme_minimal()
```

**Analysis:**

-   Categorized days into Weekday (Monday–Friday) and Weekend (Saturday–Sunday)

-   Counted accidents by `Weekend` and `Hour`

-   Visualized the distribution using a line chart

# Accident Frequency Findings: Weekday vs Weekend

**Findings:**

-   Majority of accidents occur on weekdays. Likely driven by number of drivers on the road heading to and from work.

-   Distribution of accident frequency by hour shows significant variance between weekdays and weekend

-   In order to ensure adequate emergency services, we should have more first responders on duty during weekdays from 6-10am and 3-7pm.

-   Weekends should be staffed differently to account for lower number of accidents and different frequency by hour.

# Accident Frequency: Weekend

```{r}
#| label: weekend-accidents-by-hour
#| echo: true
#| message: false
#| warning: false

# Filter data for weekends only
weekend_data <- data %>%
  filter(Day_of_week %in% c("Saturday", "Sunday"))

# Count accidents by Hour
accidents_by_hour <- weekend_data %>%
  count(Hour) %>%
  arrange(Hour)

# Plot
ggplot(accidents_by_hour, aes(x = Hour, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.8) +
  labs(x = "Hour of the Day", y = "Number of Accidents", title = "Accidents by Hour: Weekend Only") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Analysis:**

-   Accident data filtered to include only weekend days (Saturday and Sunday)

-   Count of accidents per hour of the day

-   Line chart depicting the number of accidents across each hour for weekend days.

**Findings:**

-   Accident frequency peaks at 1pm and 5pm.

-   **Afternoon (1:00 PM)**: This time sees an increase in accidents as people engage in recreational activities, errands, or social gatherings.

    **Evening (5:00 PM)**: Similar to weekdays, the evening peak occurs as people return home from various activities.

# Summary of Research and Findings: Day of the Week Analysis

-   Weekdays (Monday–Friday):

```         
-   Accident Frequency: Higher accident counts are observed on weekdays.

-   Peak Days: Fridays report the highest number of accidents.
```

-   Weekends (Saturday–Sunday):

```         
-   Accident Frequency: Lower accident counts compared to weekdays.

-   Peak Days: Saturdays and Sundays have fewer accidents overall.
```

# Summary of Research and Findings: Time of Day Analysis

-   Weekdays:

```         
-   Morning Peak (8:00 AM): Increased accidents due to morning commutes.

    Evening Peak (5:00 PM): Higher accident frequency as individuals return home from work.
```

-   Weekends:

```         
-   Afternoon Peak (1:00 PM): Rise in accidents as people engage in recreational activities.

-   Evening Peak (5:00 PM): Similar to weekdays, accidents increase as individuals return from daytime activities.
```

# Summary of Research and Findings: Final Insights and Observations

-   Weekday Accidents: The higher frequency of accidents on weekdays, particularly on Fridays, may be attributed to increased traffic volume during work commutes.

-   Weekend Accidents: The afternoon peak on weekends suggests that recreational and social activities contribute to higher accident frequencies during these times.

-   Evening Patterns: The consistent evening peak across both weekdays and weekends indicates that returning home from various activities is a common factor influencing accident frequency.
